Combining living and fossil taxa into phylogenies: the missing data issue

### 2015-08-06 Manuscript/MPESubmission/Revision/response.tex

\end{enumerate}

\section{Reviewer 2:}
Note that we split the second comment from reviewer 2 in three different comments (comments 2,3 and 4 below) to improve the readability of our response to the reviewers.

% -----------------
% Comment 1
% -----------------
\begin{enumerate}
\item{\textcolor{blue}{\textbf{Compare the trees to the "true" tree:} I don't see how the "best" tree approach is helpful in this case. It adds an additional layer of complexity that can't be accounted for via simulation. The point of simulations is to compare empirical realities to the generating process, to see how much results diverge. I think the most helpful and straightforward thing to do is eliminate the "best" tree part of the MS, and compare the inferred trees directly to the real topology to quantify variance.}}

A similar point/suggestion was also raised by reviewer 1 (see Reviewer 1 General comment number 6 and specific comment number 25).
We understand the concerns of the reviewer but, again, we do not agree that comparing the "missing-data" trees to the "best" trees adds a layer of complexity to our simulations.
In fact, we argue that comparing the "missing-data" trees to the "true" tree actually adds a layer of complexity to the analysis.
The problem we are tackling in this manuscript is to understand "what is the effect of our three parameters on recovering topology ?"
To answer this question, we have to add a layer between the data (i.e. the "complete" or "missing-data" matrices) and the actual topology which is the tree inference.
We test the effect of this layer by using different tree inference methods.

% NC: Note: problematic does not mean what you think it does... Something can be problematic (it's an adjective not a noun).

Because we use a simulation approach to test this problem, we have to start from a random starting point (the "true" tree) and then do the following steps: (1) generate and remove the data (2) infer the topologies and (3) compare the topologies.
Note that in this schematic representation, the "true" tree is the step (0) of our analysis and involves a series of algorithms and software (i.e. seqGen\{phyclust\}, rTraitDisc\{ape\}, RAxML, MrBayes).
To compare the "true" tree's topology to the "missing-data" topologies will be equivalent of exploring the effect of our three parameters on recovering topology through 3 layers (1- generate the data; 2- infer the topologies; and 3- compare the topologies).
By comparing the "best" tree's topology to the "missing-data" ones, we explore the effect of our three parameters through only 2 layers (2- infer the topologies; and 3- compare the topologies).
Note that we explore the effect of these 2 layers by using two different tree inference methods (effectively 4 since we also use the Maximum Likelihood bootstraps trees and the Bayesian posterior trees) and 2 different Normalised topological metrics (Robinson-Foulds and Triplets).
Using as main results the comparisons between the "true" and the "missing-data" trees would also need the exploration of the effect of the cascade of algorithms and softwares (maybe by modifying the algorithms used to generate the "true" tree and the matrices?).

% NC: This argument makes no sense. The point the reviewer is making is that we go:
% True tree > Best tree > comparisons when we could just go True tree > comparisons. So our approach *is* more complex than their suggestion. Instead the justification here is the same as used for reviewer one - that in real life scenarios we never know the true tree and we want it to be as close to real life as possible. BUT that we have tested their idea and find it makes no difference but all the results are in the appendix.

However, we have added analyses with comparisons between the "true" tree and the "missing-data" trees in the Appendix C (Figures C6 and C7).
It is interesting to note that the algorithms and software as well as probably the size of our matrices are actually not ideal for the recovery of the "true" tree is rather low and scales down the our results (see figure @).% NC: HUH????
However, even though these comparisons are worse than comparing the best tree to the missing data trees, it is encouraging to see that the patterns in the results are conserved (i.e. Bayesian consensus trees $>$ Maximum Likelihood trees $>$ both bootstraps trees and Bayesian posterior trees).

@@

% -----------------
% Comment 2
% -----------------
\item{\textcolor{blue}{\textbf{Better explanation of the missing data:} I found the explanation of the missing data strategy very difficult to follow. First, they oddly use the verb "to infer" for matrices; are they inferring matrices? Second, it's not clear anywhere that the authors actually removed the molecular data from the fossil taxa, though I assume they did. I hope the authors can re-write in simpler language ("explain like I'm 5") for easier comprehension.}}

The lines 180 and 181 states "All the molecular information for fossil taxa was replaced by missing data ("?")" in the former manuscript. We changed the verb "infer" (see specific comment number 16 from reviewer 1).

We assume the reviewer is referring to section "2.2 Removing data". We therefore modified the whole section to improve its clarity and simplicity as follows:

"To explore the effect of missing morphological data on topological recovery, we removed various amounts of the “complete” matrix to obtain matrices with missing morphological data. Hereafter, we call these matrices with missing morphological data the “missing-data” matrices.
% NC: add clarification here about molecular data
Note that the amount of molecular data remained constant throughout our simulations - XX in living taxa and no molecular data for fossil taxa (see above).
% NC: Is this correct?
 We removed morphological data using three data incompleteness parameters:
1. The proportion of missing living taxa ($M_L$). This first missing-data parameter corresponds to the proportion of living taxa with no morphological data. It represents the number of living taxa that are present in the matrix but have only molecular data available. This reflects the fact that, because of the increasing ease of collecting molecular data, morphological data for living species is rarely collected (Guillerme and Cooper, 2015). Therefore, many living species will have only molecular data available. In practice, we removed all the morphological data from randomly chosen living taxa with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of living taxa with no morphological data.
2. The proportion of missing data in the fossil record ($M_F$). This missing data parameter represents the completeness of the fossil record. Due to preservation biases, missing data for fossil taxa are common (Sansom and Wills, 2013). In practice, we randomly removed a proportion of data among the fossil taxa with five different proportions: 0\%, 10\%, 25\%, 50\% or 75\% of overall missing data for the fossil taxa.
3. The number of morphological characters for both living and fossil taxa ($N_C$). This parameter is not a missing data parameter \textit{per se} but rather an indication of the size of the matrix. Any morphological matrix of any size has indeterminate missing data, given that the total number of characters is undefined, but presumably large. Therefore, this parameter corresponds to the overall number of characters available for both living and fossil taxa. In practice, we randomly removed whole characters from the morphological matrix reducing it to: 100, 90,75, 50 or 25 characters.
Note that these levels are equivalent to the two other parameters (i.e. 0\%, 10\%, 25\%, 50\% or 75\% of “missing” morphological characters).

In practice, each parameter represents a different way of removing data from the morphological part of the matrix: $M_L$ removes entire rows from the living data; $M_F$ removes cells from the fossil data; and $N_C$ removes columns across both living and fossil data. Note that $M_L$  and $M_F$ differ not only because of the region of the matrix affected: for $M_L$ all the morphological data of a percentage of living taxa are removed, whereas for $M_F$ a percentage of the data are removed at random from across the whole of the morphological matrix for fossil taxa.

@@

We created matrices using all parameter combinations resulting in 125 ($5^3$) “missing-data” matrices. Note that one of these combinations ($M_L$=0\%; $M_F$=0\% and $N_C$=100) has no missing data so is equivalent to the “complete” matrix, thus we have one effectively complete matrix in our 125 “missing-data” matrices. In practice, we first removed the data following the two missing data parameters $M_L$ and $M_F$ and then removed data following the $N_C$ parameters. To avoid avoid matrices containing taxa without any data (morphological or molecular), we repeated the random deletion until the matrices contained at least 5\% of data for any taxon. Note that the living taxa always had at least 90\% of data (the 1000 molecular characters)." lines 222-267@@@

% NC: NC_C here again not M_C...

% -----------------
% Comment 3
% -----------------
\item{\textcolor{blue}{\textbf{$M_C$ is not missing data per se:} If Mc is just removing columns, then they are shrinking the total matrix. This is not missing data per se. They are instead comparing matrices of different sizes, and the quantity is "data present." Any matrix of any size has indeterminate missing data, given that the total number of characters is undefined, but presumably large. A matrix with 10,000 characters is tiny if 10,000,000 could be scored. This is semantic.}}

% TG: Ok i've changed it every where but personally I don't like it. It makes it even more messy and hard to read. However, because I didn't answered the first and last comment of this reviewer, I felt that at least 2/4 is a minimum.
We agree with this comment and hope it will help to the global comprehension of the manuscript. We replaced the $M_C$ (missing characters) parameter by $N_C$ (number of characters with five states: 100, 90, 75, 50 and 25) in the revised manuscript (see reviewer 2 comment 2 above), the appendices and the figures. We also slightly modified some explanations related to the change of missing characters to number of characters throughout the manuscript.

% NC: Ah OK. You haven't corrected this everywhere in these comments I don't think...

@@

% -----------------
% Comment 4
% -----------------
\item{\textcolor{blue}{\textbf{Specific variation of data per fossil:} More importantly, it doesn't seem like missing data varied among fossil taxa, unless I'm misreading. What we really want to know is what happens when some fossil taxa are fragmentary, as this is the most common empirical situation. Thus, there are morphological characters scored for some living taxa, some fossil taxa, and missing from others. It's unclear if this was simulated. It would be good if the authors could detail the case where, say, all living taxa had morphological data (0 Ml), the matrix was large (0 Mc), and Mf varied *among* fossil taxa, such that some fossils had 100\% coverage, and varied down to, say, 5\%. That's the crucial case here.}}

In this study, we applied a similar approach to what the reviewer 2 suggests (i.e. no missing data for living taxa, and maximal number of characters but various amount of missing data for the fossil taxa). 
These results are present in figure 4, central row.
However, we think we may have not explained ourselves entirely clearly. The amount of changes among the fossil taxa was not set to any specific amount of missing data per fossil (e.g. like the "artificial fossil" approach in Pattinson et al 2014, Syst. Biol.) but we rather treated the missing data in the "whole" fossil record.
Because our fossil taxa and their characters where randomly simulated, we didn't had any \textit{a priori} assumption on how to remove the data.
We therefore decided to randomly remove a certain amount of data among all the fossils.
For example, when removing 75\% of the data for the fossil taxa ($M_F$=75\%) different scenarios could happen, ranging from a "homogeneous" incompleteness data distribution (each specific fossil taxa had 75\% of missing data) to a "heterogeneous" incompleteness data distribution  (80\% of the specific fossil taxa had 5\% available data and 21\% of the specific fossil taxa had 100\% available data).
In empirical matrices, the incompleteness data distribution probably lies in between both models.

% NC: OK - I'd add a line somewhere in the manuscript to make this clear. something like:

%"Note that 55\% missing data for fossil taxa does not mean that each fossil is missing 50\% of it's morphological data. Instead this 50\% refers to the whole matrix. Some fossils may reatin 100% of their data and others may lose all their data at this parameter value"

Also, a second argument for using this slightly simplistic way to deal with the missing data in the fossil record follows our main recommendation from this study: the best way to improve the quality of Total Evidence trees is to have a maximum number of living species with available morphological data.
In fact lines 584 to 589 in the first version of the manuscript (line 669-674@@@ in the revised version), we argue that even if the effect of the $M_F$ parameter is important, there is unfortunately not much we can do about it since the completeness of this part of the matrix solely depends on exceptional palaeontological discoveries and excavation effort.
We feel that these parameters, however crucial for our understanding of macroevolution, are unfortunately not predicable and a recommendation such as "we need more fossil data" is less practical than our main recommendation: "we need more living species data".

% NC: Not sure the second paragraph is needed. The reviewer is saying that they want to know what happens if living taxa have all their data but fossils have variable amounts. And I think this is what you've done for some parameter combinations. Just need to apologise for not being clear and clarify in the text how data were missing from fossils. Doesn't need so much waffle and defensiveness - they are asking you to do what you've already done! :)

### 2015-08-06 Manuscript/MPESubmission/Revision/response.tex

% -----------------
% Comment 5
% -----------------
\item{\textcolor{blue}{\textbf{Abbreviation:} One suggestion is to consider abbreviating Maximum Likelihood and Bayesian Inference as ML and BI respectively early in the methods and use throughout.}}

We understand the reviewer's concern.
Nonetheless, we would prefer not abbreviate these terms. Although they might make sentences where Maximum Likelihood and/or Bayesian Inference are stated more concise, we think spelling out the terms makes our paper clearer and does not ask the reader to look back for the actual meaning of the abbreviation (Berlin 2013; Radiology; \url{http://pubs.rsna.org/doi/abs/10.1148/radiol.12121776}).
However, we can change this at a later stage if necessary.

% -----------------
% Comment 6
% -----------------
\item{\textcolor{blue}{\textbf{Methodology - True tree:} I suggest it warrants comparing at least the 'best' tree to the 'true' tree.}}

We have now compared both the ``best'' tree and the ``missing-data'' trees to the ``true'' tree as suggested by both reviewers.
The results are available in appendix B; figure B1 and appendix C; figures C6 and C7.
The ability to recover the ``true'' tree is poor in both methods and using both metrics (Maximum Likelihood 50\% CI approximatively between 0.2 and 0.3 Normalised Robinson-Foulds and Triplets metric; Bayesian consensus tree 50\% CI approximatively between 0.3 and 0.4 Normalised Robinson-Foulds metric and between 0 and 0.2 Normalised Triplets metric).
This can be related to the size of our matrices or to the ability of the algorithms to generate the ``complete matrix'' but also to our decision to put our minimum median bootstrap acceptance at 50 rather than higher values (line @@@ in the former manuscript and @@@ in our revision).
When comparing the ``true'' tree to the ``missing data'' trees, the results are \textit{scaled down} due to the low scores of the ``true'' \textit{vs.} ``best'' trees scores.
However, we can observe the same patterns between the methods as when using the ``best'' tree (i.e. Bayesian always outperforms ML in clade conservation but does poorly in wild-card taxa placement).
An interesting feature is that the Normalised Robinson-Foulds metrics actually \textit{increases} with missing data.
This is due to the fact that the difference between the ``best'' and the ``true'' tree is high making trees with enough data more different.
When data is missing, however, more clades are unresolved and thus closer to clades from the ``true'' trees.

% NC: Worth describing the results here.

@@

Although though we agree with the comments of both reviewers that knowing the difference between the true and best/missing-data trees is important, we do not think that these results warrant inclusion in the main text.
In these simulations, we had to make a choice between (1) simulating larger matrices but with fewer chains (i.e. less variance) or (2) smaller matrices with more variance.
We deliberately made the second choice because our main question was "what is the effect of our parameters on recovering topology".
To answer this question, we maintain that comparing the best tree to the missing-data trees is the best approach for two reasons:
\begin{enumerate}
\item{The true tree is always unknown:} as we state in the manuscript: "In real world situations, the ``true'' tree is never available" (line 242 in the first manuscript and @@@ in the revision).
\item{The true tree is just the starting point of the simulations:} we use the true tree only as a starting point for our simulations.
Once generated, the true tree is fed to a cascade of algorithms and software (i.e. seqGen\{phyclust\}, rTraitDisc\{ape\}, RAxML, MrBayes) described in section 2.1 and 2.3 to get the best tree and the missing-data trees.
Comparing the latter to the true tree will answer our question "what is the effect of our missing-data parameters and the cascade of algorithms and software on recovering the true topology?".
\end{enumerate}
Therefore we added a second justification to the main text (see below this response) and added these analysis (true vs. best/missing-data trees) to the supplementary for readers that might be interested by this specific problem.

We added the following sentence to the main text to clarify and link to the supplementary analysis:
% "In practice, the difference between the ``true'' and the ``best'' tree represents the effect of our parameters and R packages choices %NC: What dod you mean by R packages choices? Do you mean the way the trees were simulated?
%  (i.e. diversitree, phyclust and ape) as well as the phylogenetic methods used (see appendix B).
% The aim of this study, however, is to look at the effect the missing data parameters described above on topological recovery." line 276-280 @@@.

% TG: changed to:
"We compare ``best'' trees to ``missing data'' trees but could also compare ``true'' trees to the ``missing data'' trees.
In practice, the difference between the ``best'' trees and the ``missing data'' trees represents the effect of our missing data parameters and of the phylogenetic methods used to infer the ``missing data'' trees.
The difference between the ``true'' and the ``missing data'' tree, however, represents the effect of our parameters used to generate the ``true'' tree and the algorithms used to generated the ``complete'' matrix as well as the effect of our missing data parameters and the phylogenetic methods used.
Because the main aim of this study, is to look at the effect the missing data parameters on topological recovery, we chose to represent only the comparisons between the ``best'' trees to ``missing data'' trees.
The results of the comparisons of the ``true'' and the ``missing data'' tree are available in appendix B.
Note that these results makes little difference to our overall results." line 276-280 @@@.

See also appendix B section "Differences between the ``true'' and the ``best'' trees."
% NC: Needs a little more explanation, and a line about what analysis you did in appendix b. So something along the lines of "We compare best trees to missing data trees, but could also compare true trees to the missing data trees. We do not do this because... However, results of analyses comparing the two can be found in appendix B. This makes little difference to our overall results." That might need to be better integrated with the paragraph(s) before.

@@

% -----------------
% Comment 7
% -----------------
\item{\textcolor{blue}{\textbf{Methodology - Starting tree:} The authors should at least include some analysis in which no starting tree is given, the max. likelihood 'best' tree is given, and / or the 'best' tree is given but with some perturbation.
I fear that one of the main results may be biased due to this one methodological detail - a strong prior on tree topology.
The justification given was 'to speed up the Bayesian analysis' which does not seem adequate, given these datasets are small.}}

We performed an extra analysis on a subset of five trees with the three missing data parameters at 0\%, 10\%, 25\%, 50\% and 75\% for 20 chains. 
We compared the topology of these trees in Bayesian inference when using the true tree as a starting tree or using a random tree (default in MrBayes) as a starting tree.
% NC: Yes but the reviewer is asking for something different - they ask you to have no starting tree. If this is impossible in MrBayes then say so, and say you used a random tree instead. TG: Maybe there's a phylogeny confusion here. In both ML or Bayesian, you need to start from a tree, then calculate the likelihood of all the parameters, then change the tree, calculate again, etc...
We found no difference in topologies between both the Bayesian inferences starting with the true tree or a random tree.
Therefore, we do not believe our results are merely the result of using a specific starting tree topology. %strong priors on the tree topology. TG: That's the big difference that the PLoS ONE reviewer also didn't got. A starting tree is not a Bayesian prior (in P(A|B) = P(B|A)*P(A)/P(B; P(A) is the prior). It's just a user fixed starting point for the MCMC walk rather than a random one.
% NC: The reviewer also says that saving time is not a good justification. It'd be worth explaining here how long it actually takes even though these datasets are small.
%
% TO DISCUSS @@@
% TG: about that: it using a random tree or the true tree doesn't change the timing significantly. The whole idea was suggested by Rob or Gavin (don't remember exactly) at lunch time when I was visiting Gavin in Sheffield. To speed up the search in the tree space they suggested I already start from a sensible tree in the tree space, that was meant to speed up the analysis.
%
These results are available in the supplementary materials Appendix A (figure A2 and tables A1 and A2) and we also added a note to the manuscript:

@@

Concerning the last question of this comment: "How then in comparisons of the distributions are they always so low (Table 1 max prob. of overlapping distributions = 0.11)?" when looking at all parameters separately (figure 4), there is no clear overlap between the Bayesian consensus trees and the Bayesian posterior trees (median Bhattacharyya Coefficient of 0.01; table 1).
The maximum observed Bhattacharyya Coefficient (0.11; table 1) is probably due to the combination of parameters: when all the parameters have many missing data, the probability of overlap between trees was slightly higher (0.11) but still really low.
For the Normalised Triplets distance however, the overlap is higher (median Bhattacharyya Coefficient of 0.59; table 1) as illustrated in figure 4 (i.e. much more overlap between the distributions).

% TG: I'm confused for the second part of the question:
% TG: In figure 4 and in Table 1, if you look at the Bayesian consensus (grey) trees and the Bayesian posterior trees (orange) for RF, they do not really overlap. Ever. The 0.11 probability must come from the parameters combinations that, when is really shitty must show a little bit of overlap. But for each parameter one by one: nope. No overlap.
% TG: On the other side, in figure 4 and Table 1, for Bayesian consensus and Bayesian posterior for Tr, the do overlap much more. And that's what's in the Table 1 (min 0.21; median 0.56; and max 0.87).
% TG: Did the reviewer just miss read the labels of the Table 1 or am I miles away from what he means?

% NC: It's more likely an issue with reading figure 4. Check your figure legends. I think they might be incorrect. Then the reviewer might be seeing overlap of the wrong two distributions... But if they do overlap all the way to 25% then it is weird that the highest BC is 0.11 isn't it?
% TO CHECK @@@
% TG: yes definitely. Just doubled check, it might be because the colors where miss labelled and the reviewer miss read the graph (my fault).

@@

% -----------------
% Comment 10
% -----------------
\item{\textcolor{blue}{\textbf{Results - Bayesian Triplets:} Another important result that is not mentioned is that using the wildcard taxon metric (Tr), Bayesian consensus trees do quite poorly with missing data.
Values in Figures 4 and 5 clearly drop below zero, although negative values are not illustrated on the axis, and this indicates that the tree topologies are no better than, or even worse than, random trees.
It is especially true when fossils are missing data, and that is going to be an important finding with practical application but must be addressed. }}

We modified the figures 4 and 5 to show Normalised Triplets metric values below 0 and added a dashed line at 0 to underline the difference in scale between both metrics.
As mentioned above (reviewer 1 comment 9), this is due to the fact that the Bayesian consensus tree will be less and less resolved when missing data increases.
This has two related effects.
The first one being to retain "good" clade conservation (i.e. avoiding false-positive nodes by collapsing unsupported ones).
And the second one is that when these clades are collapsed, it reduces the number of available triplets and thus increases the wild-card taxa score (Normalised Triplets metric).
In fact, when a clade ((a;b);c); is collapsed to (a;b;c), the whole clade is conserved (a, b and c are still in a common clade) however they also all become wild-card taxa: the position of a regarding b and c is now incorrect.

We added a note on this in the section about the effects of the tree inference method:

"The Bayesian consensus trees, however, perform badly for the Normalised Triplets metric: some parameter combinations, especially when the $M_F$ parameter reaches 75\% missing data, lead to negative values (figure 5).
A Normalised Triplets metric value below 0 means that the placement of some taxa is worse than expected by just randomly placing this taxon in the tree.
This can be interpreted by the absence of comparable triplets between some of the ``missing data" trees and ``best'' trees.
Even if clades are conserved (figure 5), the resolution within them can be poor to non-existent when a lot of data is missing (i.e. 75\%).
In such cases, the fossil taxa are likely to be placed in any of the clades that share the most characters.
These results are in agreement with previous studies that have showed that missing data can cause problems for recovering ``correct'' topologies, especially for small matrices of 100 characters (Wiens 2003).
It is important to note, however, that this effect can be reduced by increasing the number of characters (Wiens 2003)." lines 632-644. @@@

@@

% -----------------
% Comment 11
% -----------------
\item{\textcolor{blue}{\textbf{Conclusion:} The authors should clarify how their results compare to those of previous studies more directly to emphasize the nuances that make this study unique. }}

We rewrote the conclusion to deal with this comment and the specific comments 13, 52 (see comment 52 for details on the choice of citing Arcila et al 2015.) and 56.
% NC: Is it Arcial or Arcila? TG: Dahiana Arcila

"Previous studies have explored the effect of missing morphological data when combining living and fossil species in phylogenies but where limited by their empirical approach or their aspects of missing data (Wiens et al 2005; Pattinson et al 2014). % NC: I still don't get these limitations. How are they are issue? What aspects?

% TO DO: develop the limitations @@@

This study proposes a more in depth approach where missing data is generated on simulated data and along three clear missing-data parameters ($M_{L}$, $M_{F}$ or $N_{C}$).
This allowed us to confirm previous results that missing data can be especially problematic in small matrices (Wiens 2003) but also revealed the crucial importance of coding morphological data for living species.
Missing data in Total Evidence matrices is not a problem for recovering the ``best'' tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters.
When missing data increases in any of our missing data parameters ($M_{L}$, $M_{F}$ or $N_{C}$), it reduces support for the placement of fossil taxa and increases the displacement of wildcard taxa.
Therefore we advise that one should focus on coding morphological characters for a large number of the living taxa present in the matrix (i.e. 50\%) for accurately combining both living and fossil species in phylogenies.
Doing so grants a good overlap of morphological characters among living and fossil taxa, allowing the fossil taxa to be positioned relative to the living taxa based on their actual shared derived characters rather than simply on available data."

Additionally, the topology of the Bayesian consensus trees, regardless of the amount of missing data, were always closer to the ``best'' tree topology than the Maximum Likelihood trees.
This has also been observed in empirical data (e.g. Arcila et al., 2015) where Maximum Likelihood trees inferred from a Total Evidence matrix were less supported than the Bayesian consensus tree.
This might have an important impact on estimating topologies in the Total Evidence framework, because previous studies had to rely either on molecular scaffolds (e.g. Slater, 2013), taxonomic constraints (e.g. Slater, 2013; Beck and Lee, 2014) or even by fixing the topology (e.g. Ronquist et al., 2012a).
Therefore, we suggest extracting such topological backbones from the Bayesian consensus tree if needed.
% TG: maybe add in a non tip-dating analysis?

@@

% -----------------
% Specific Comment 50
% -----------------
\item{\textcolor{blue}{\textbf{line 612-613:} too obvious and not especially helpful. We should aim to have as little missing data as possible…more helpful, and something especially discovered in this study, is concentrating on having overlapping morphological characters between the living and fossil taxa to be able to place the living taxa with strong support via large datasets (combined data) and being able to position the fossils based on their shared derived characters with living taxa. This is an important point to make clearer. }}

Agreed. See reviewer 1 comment 11 for the changes to the conclusion.

% -----------------
% Specific Comment 51
% -----------------
\item{\textcolor{blue}{\textbf{line 613-615:} rephrase - sentence is choppy with all the commas. }}

We removed this sentence and replaced it by the two new sentences outlined above (reviewer 1 comment 11).

@@

%NC: Ummm but specific comment 50 refers to comment 11. Should this also refer to comment 11?

% -----------------
% Specific Comment 52
% -----------------
\item{\textcolor{blue}{\textbf{line 615-618:} It seems more important to me that the authors emphasize what implications it may have for previous and future studies that Bayesian analyses outperform max. likelihood, other than to re-hash that the authors feel the Bayesian consensus should be used to fix topologies for other downstream analyses.
For example, can the authors cite some combined analyses that used only likelihood analyses and came up with controversial results?
Or what about other studies that have done combined analyses with different optimality criteria and found different topologies?
What about implications for studies that have used molecular scaffold techniques to fix the topology of living taxa based on analyses of only DNA analyses and then fit the fossils into this fixed tree?
Also 'further' is misspelled.  }}

Unfortunately, we did not find many studies formally testing the important questions asked by the reviewer. However, there are a few empirical analyses that contain results supporting our claims.
Firstly, for the question about the topological differences between Maximum Likelihood and the Bayesian consensus tree, in Arcila et al 2015 (MPE)'s supplementary analysis, the authors compare the topology inferred from a Total Evidence matrix in both Likelihood (RAxML) and Bayesian (MrBayes) and the topology of the Bayesian consensus tree is much better supported than the one of the best ML tree. Specifically, the ML tree has 7/28 nodes have a "fair" bootstrap support ($>$ 75) and 3/28 with a good support ($>$ 90) as the Bayesian consensus tree has 17/25 nodes with a "fair" posterior probability ($>$80) and 14/25 with a good posterior probability ($>$95). We therefore used this reference to strengthen our claim of the superiority of the Bayesian consensus tree over the Maximum Likelihood one.
Secondly, for the question about molecular scaffolds, from the few studies clearly specifying their topological constraint method, only Beck and Lee 2013 (Proceedings B) seems to compare an analysis with and without taxonomic constraint on divergence time and rate estimation (the constraint seems to have no effect) but the branch length aspect is not covered in our study.

See our response to reviewer 1's comment 11 above for changes in the revised manuscript.
